\section{Methodology}
% =======================================================================
\subsection{Data Collection}
% -----------------------------------------------------------------------
\subsubsection{Data Sources}
English-speaking users who discuss finance and investing on online social media platforms in text form do so on three major platforms: Reddit, StockTwits, and Twitter. Reddit is a SNS on which users can create their own communities (``subreddits'') that focus on a specific topic. For example, users have created the subreddits ``Investing'' and ``StockMarket'' to discuss long-term investments and the subreddit ``WallStreetBets'' for posts about high-risk short-term gambles in the market. However, posts on Reddit tend to be much longer than posts on Twitter or StockTwits. Their length would require them to be analyzed on a paragraph- or sentence basis. Since research on SA is mostly focused on document-level analysis, we will not use Reddit posts for this work. The decision between StockTwits and Twitter is harder: posts on both platforms are similar in length and share the usage of cashtags (a ``\$'' sign followed by a ticker symbol) for identifying stocks. We decide to obtain data from Twitter rather than StockTwits for the following reasons:
\begin{enumerate}[noitemsep]
	\item The post volume on Twitter is higher than it is on StockTwits.
	\item The few data sets for SA on financial SM posts that exist use data from StockTwits, hence publishing a data set of Tweets provides more value to the research community.
	\item By using Twitter data for our experiments we can answer RQ4 and compare performances between models and data sources.
\end{enumerate}



% -----------------------------------------------------------------------
\subsubsection{Sampling}
\begin{itemize}[noitemsep]
	\item Twitter: convenience sampling of last $7d$
\end{itemize}


% =======================================================================
\subsection{Data Labelling}

% -----------------------------------------------------------------------
\subsubsection{Task Definition}
\begin{itemize}[noitemsep]
	\item why pos/neg/neu?
	\item codebook: UoA = document! Locking in profit = positive, even if that means selling a stock. This here is not aspect(i.e. stock)-based, but text sentiment in a tweet
\end{itemize}

\input{assets/tables/static/codebook.tex}


% -----------------------------------------------------------------------
\subsubsection{Data Quality Assessment}  % TODO: wording
\begin{itemize}[noitemsep]
	\item de-dupe
	\item length/cashtag filter
	\item potentially: spam classifier model?
	\item mention: no active learning (\url{https://blog.fastforwardlabs.com/2019/04/02/a-guide-to-learning-with-limited-labeled-data.html})
\end{itemize}




% =======================================================================
\subsection{Data Preprocessing}
\label{section-data-preprocessing}
% -----------------------------------------------------------------------
\subsubsection{Tokenization}
\begin{itemize}[noitemsep]
	\item tokenization
\end{itemize}

% -----------------------------------------------------------------------
\subsubsection{Representation of Text Data}
\begin{itemize}[noitemsep]
	\item vectorizer (TF-IDF?)
	\item Embeddings?
\end{itemize}






% =======================================================================
\subsection{Experiment Design}

%todo: this needs a section for descriptive experiments like bias check and sample tweets demo

% -----------------------------------------------------------------------
\subsubsection{Performance Evaluation}
\begin{itemize}[noitemsep]
	\item cross validation
	\item metrics
\end{itemize}



% -----------------------------------------------------------------------
\subsubsection{Selection of Models}


Models to build:

\input{assets/tables/static/models_to_build.tex}



\begin{itemize}[noitemsep]
	\item model types
	\item hyperparameters
\end{itemize}

% -----------------------------------------------------------------------
\subsubsection{Hyperparameters/Search Space?}


\textbf{Finally:} Show a flowchart of everything?






